---
title: "Data Wrangling in R"
subtitle: "How to Check & Clean Missing Values / NA using R"
author: "Kaixin Wang"
date: "OSCR - Summer Session C 2019"
output: rmarkdown::github_document
---

**Note**: Some contents are based on the book *An introduction to data cleaning with R*, by Edwin de Jonge and  Mark van der Loo.

1. To import the input dataset:

```{r, message = F}
library(readr)
input = read_csv("person_info.csv")
input
```

2. To find the location of the NAs:

```{r}
idx = which(is.na(input))   # find the position of the NAs
idx   
```

3. To find the location where the character "NA", "na", "N/A", or "n/a" are:

```{r}
# NAs = c("na", "n/a", "NA", "N/A")
nas = which(input$car_1 == "na" | input$car_1 == "n/a" | input$car_1 == "NA" | input$car_1 == "N/A")
nas
```

4. To find whether one row contains a missing value:

```{r}
complete.cases(input$car_1)
```

5. To impute the missing values with NA:

```{r}
input$car_1[nas] <- NA   # impute with NAs
input$car_1
```

6. To omit row with missing value(s): 

```{r}
na.omit(input)   # remove rows that contain any NA
```

7. Typical approaches to handle missing values:

- impute with `NA`
- impute with the median/mean of the data 

8. To ignore missing values in calculations:

```{r}
# example:
vars <- c(10, 20, NA, 25, 5, 8, NA, 11)
mean(vars)   # NA introduced
mean(vars, na.rm = T)  # ignore the NAs and compute the summary statistics
```

