---
title: "Data Wrangling in R"
subtitle: "How to Check & Convert Data Types using R"
author: "Kaixin Wang"
date: "OSCR - Summer Session C 2019"
output: rmarkdown::github_document
---

**Note**: Some contents are based on the book *An introduction to data cleaning with R*, by Edwin de Jonge and  Mark van der Loo.

1. To import the input dataset:

```{r, message = F}
library(readr)
input = read_csv("person_info.csv")
input
```

2. To check the data type of the object:


(1) `str`() function

```{r}
str(input)  # data structure of the object
```

(2) `summary`() function:

```{r}
summary(input)  # summary statistics of the object
```

(3) `class`() and `typeof`() functions

```{r}
class(input$first_name)  # class of the object
typeof(input$first_name)  # type of the object
```

```{r}
class(input$age)  # class of the object
typeof(input$age)  # type of the object
```

3. Type coercion: 

 `as.numeric`, `as.logical`, `as.integer`, `as.double`, `as.factor`, and `as.character`
 
(1) examples:
 
```{r}
age <- input$age 
summary(age)
as.double(age)
as.logical(age)   # to a logical vector
as.character(age) # to a character vector
as.factor(age)    # to a factor
```
 
(2) `as.logical`:

- if the input is numeric: 0 is converted to FALSE, else converted to TRUE
- if the input is character: NA produced

(3) `as.factor`:

- the order of the levels is in alpha-numeric order

4. Datetime object: `lubridate` library

```{r, messgae = F}
library(lubridate)
dates <- c("15/02/2013", "15 Feb 13", "It happened on 15 02 '13")
dmy(dates)

mdy(input$birthday)
```

Note: There are similar functions for all permutations of d, m and y. Explicitly, all of the following
functions exist: `dmy myd ydm mdy dym ymd`.

5. character manipulation:

(1) string normalization

```{r}
library(stringr)
str_trim(" hello world ")
str_trim(" hello world ", side = "left")
str_trim(" hello world ", side = "right")

toupper("Hello world")
tolower("Hello World")

input$City[5]
input$City[5] = str_trim(input$City[5], side = "right")
input$City[5]
```

(2) approximate string matching

```{r}
gender <- c("M", "male ", "Female", "fem.")
grepl("m", gender)
grep("m", gender)
```
